\thetitle

### 2020-07-01 writing/introduction.tex
revisions to the introduction

Since its very beginning [cite STH], the theory of gradual typing
comes with a blame theorem. ``Well-typed [components]\footnote{The
original authors got this word wrong, and more.} can't go wrong'' has
become the slogan. The idea is that blame assignment can help the
practicing programmers find impedance mismatches between the types
they added and the remaining untyped components.

Practical gradual type systems ignore blame completely. Industrial
systems [lots of citations. footnotes] exploit types for assistance
with IDE actions and for finding typos in code. Then they remove them
and rely on the built-in safety checks of the underlying programming
language to catch problems.

This contrast raises the question whether

     blame adds value to gradual typing
     and especially the working programmer.

To obtain an answer, we have developed a method (not methodology!)  to
evaluate blame assignment systems from the gradual typing
world. Roughly speaking, the method simulates a programmer who follows
the above-cited Wadler-Findler slogan. When the run-time system
signals a violation, the programmer conjectures that the addition of 
types to untyped components turns the dynamic violation into a static
type error and thus reveals the impedance mismatch.

The paper presents the evaluation method and the results of applying
the method to three distinct blame assignment systems.

- explain the systems
- explain the method
- preview the results

- weave roadmap into these paragraphs

>>>>

From the very beginning, the theory of gradual typing has come with a blame
theorem~\cite{mf-toplas-2009, tf-dls-2006}. ``Well-typed [components]\footnote{The
original authors got this word wrong.} can't be blamed'' turned the theorem into a
slogan~\cite{wf-esop-2009}. Indeed, academic implementationsa---for example,,
Reticulated Python~\cite{vsc-dls-2019, vss-popl-2017, vksb-dls-2014} and Typed
Racket~\cite{tf-dls-2006,tf-popl-2008,tfffgksst-snapl-2017,tf-icfp-2010}---come
with sophisticated blame-assignment subsystems. Theoreticians firmly embrace the
idea that blame can help practicing programmers find impedance mismatches between
the types they added to a software system and the behavior remaining untyped
components.

Practical gradual type systems ignore blame assignment almost completely.
Industrial systems such as Flow,\footnote{flow} Hack,\footnote{hack} or
TypeScript~\cite{ts} exploit types for IDE actions and for finding typos in
code. Then these compilers remove types and rely on the built-in safety checks of
the underlying programming language to catch any problems.

The contrast between academic and industrial systems raises the question whether
\begin{quote}
 \it
 blame assignment adds any value to a gradually typed language, \\
 especially for the benefit of the working programmer.
\end{quote}
Given the prominence of blame in academia and its complete absence in
industrial systems, it comes as an even bigger surprise that the research
literature and the industrial blog world do not present or discuss any possible
answers.  Instead, language designers seem to answer this question one way or
another without any scientific justification, and even when they have a language,
they neglect to look into this question. The research community simply does not
have a method for evaluating such questions, and given the question's practical
relevance, this is a problem.

This paper's first contribution is an automated method for evaluating the
effectiveness of blame assignment strategies in the gradual typing world (see
section~\ref{sec:ideas}). Roughly speaking, the method simulates a programmer who
follows the Wadler-Findler slogan quoted above. Its implied method suggests that
programmers can find impedance-mismatch bugs by adding types to a partially typed
software system. That is, when the run-time system signals an impedance mismatch,
a good blame messages helps programmers infer where to add an additional type
specification. Since components that pass the type checker cannot cause such
impedance signals; the type checker must discover these mismatches at compile
time.  If the additional type does {\em not\/} cause a static type error, the
program is run again until it signals another violation.  Counting the number of
steps it takes to discover the mistake is a measure of the effectiveness of the
blame assignment.  The method is loosely based on the work of Lazarek et
al.~\cite{lksfd-popl-2020}.

The paper's second contribution is a set of results from applying the evaluation
method to three distinct blame assignment systems and approximately 20,000
cases\mf{Replace with real number} (see section~\ref{sec:experiment}): Vitousek et
al.'s strategy of tracking typed/untyped boundaries in Reticulated;
Tobin-Hochstadt et al.'s use of higher-order contract system~\cite{ff-icfp-2002,
mf-toplas-2009}; and the error messages from failed safety checks in the
underlying programming language.

\mf{This is made up and must be fixed for the submission.}

The results of the experiment (see section~\ref{sec:results}) validate the
conjecture behind the work of theoreticians. A good blame assignment system
greatly shortens the search for impedance mismatches between the specified types
and the behavior of untyped components. Even better, the evaluation method clearly
identifies the wrapper-based blame tracking of Typed Racket as the vastly superior
academic system. Vitousek et al. lose one more time. Of course, a detailed
discussion must acknowledge the threats to validity of these results (see
section~\ref{sec:discussion}).

Next, the paper provide an overview of the blame assignment strategies as found in
existing gradually typed language implementation (see section~\ref{sec:landscape})
and the challenges of developing a systematic evaluation method
(section~\ref{sec:challengs}).
